---
title: "Sleep, Health, and Lifestyle Dataset: Sleep Duration Analysis"
author: Justin Naidu
format: html
theme: superhero
execute:
  echo: false
  error: true
---

```{r message=FALSE}
library(tidyverse)

ss <- read_csv(here::here("data", "raw", "ss.csv"))

```


```{r}
ss_clean <- ss %>%
	rename(
		Person_ID = `Person ID`,
		Sleep_Duration = `Sleep Duration`,
		Sleep_Quality = `Quality of Sleep`,
		Physical_Activity_Level = `Physical Activity Level`,
		Stress_Level = `Stress Level`,
		BMI_Category = `BMI Category`,
		Heart_Rate = `Heart Rate`,
		Daily_Steps = `Daily Steps`,
		Sleep_Disorder = `Sleep Disorder`
	)
```

# Data Description

The data set examined in this final project report for EPI 590R is the "Sleep, Health, and Lifestyle" data set. This data set was created by Henry Shan and is accessible through the Kaggle website.The data set contains a wide range of variables related to sleep and daily routines. The data set includes variables such as gender, age, occupation, sleep duration, sleep quality, physical activity, stress level, BMI, blood pressure, heart rate, daily steps taken, and presence of sleep disorders.

<https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data>

# Table of Descriptive Statistics

```{r}
#| label: fig-table1
#| fig-cap: "This table presents the distribution of key characteristics among female (N = 1,841) and male (N = 1,891) participants. Values are expressed as median (Q1, Q3) or n (%). The characteristics include sleep duration, physical activity level, BMI category, heart rate, daily steps, and presence of sleep disorders. Notable findings include similar median sleep durations and physical activity levels between genders, but significant differences in BMI categories and the prevalence of sleep disorders, particularly sleep apnea, which was more common among females."

library(gtsummary)

table1 <- tbl_summary(
  ss_clean,
  by = Gender,
  include = c(Gender, Sleep_Duration, Physical_Activity_Level, 
  						BMI_Category, Heart_Rate, Daily_Steps, Sleep_Disorder),
  label = list(
  	Sleep_Duration ~ "Sleep Duration",
  	Physical_Activity_Level ~ "Physical Activity Level",
  	BMI_Category ~ "BMI Category",
  	Heart_Rate ~ "Heart Rate",
  	Daily_Steps ~ "Daily Steps",
  	Sleep_Disorder ~ "Sleep Disorder"
  ))

table1
```


# Univariate Regression

```{r}
#| label: univar-reg
#| fig-cap: "Females, on average, have 0.19 hours less sleep duration compared to males (p-value < 0.021). Higher stress levels are linked to shorter sleep duration (p-value < 0.001). Overweight individuals have 0.63 hours less sleep duration (p-value < 0.001)."
	
univar_reg <- tbl_uvregression(
  data = ss_clean,
  y = Sleep_Duration,
  include = c(
    Gender, Sleep_Duration, Sleep_Quality, Stress_Level, 
    BMI_Category, Sleep_Disorder),
  method = lm
)

univar_reg

```

# Figures
```{r}
#| label: fig-boxplot
#| fig-cap: "In this boxplot, comparing both males and females in the dataset, the median sleep duration for females (represented by the red boxplot) is approximately 8 hours. The median sleep duration for males (represented by the blue boxplot) is around 7 hours. Females tend to have longer sleep durations."

library(ggplot2)


# Boxplot by Gender
boxplot <- ggplot(ss_clean, aes(x = Gender, y = Sleep_Duration, fill = Gender)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Sleep Duration by Gender",
    x = "Gender",
    y = "Sleep Duration"
  )

boxplot
```

@fig-table1 contains a table that illustrates the baseline characteristics of study participants by gender and @fig-boxplot contains a boxplot that compares the sleep duration between males and females.

In our dataset, `ss_clean`, there are `r nrow(ss_clean)` participants. Among them, `r sum(ss_clean$Gender == "Female")` are female and `r sum(ss_clean$Gender == "Male")` are male. The average `Sleep_Duration` for male participants is `r gtsummary::inline_text(table1, variable="Sleep_Duration", column="Male")` hours, while for female participants it is `r gtsummary::inline_text(table1, variable="Sleep_Duration", column="Female")` hours. 

Additionally, the standard deviation of `Sleep_Duration` across all participants is `r round(sd(ss_clean$Sleep_Duration, na.rm = TRUE), 2)` hours.


